---
title: "How Fast Can You Type A Million Letters?"
subtitle: 'Prepared for FiveThirtyEight Riddler Classic April 27, 2018'
output:
  html_document:
    keep_md: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nloptr)
```

### How I Solved It
I used an iterrative optimization approach to solve this problem.  I found a non-linear constrained optimization library in R `library(nloptr)` that implements the NLopt library. I set up the optimizer to minimize the sum of the times spent on various methods (including the delays specified in the problem), subject to the constraint that one million i characters are typed.  First I tried this without copy-paste cycles, and then progressively added additional copy-paste cycles.  It took seven cycles until the decrease in total time leveled off.  I chose the solution with six copy-paste cycles for a total time of 41 seconds.   Please note that this optimizer doesn't do integer optimization, so I hand tweaked (rounding up or down) the solutions to get whole numbers.  I used the sign() function to model the fixed costs of operations such as copy/paste cycles. For optimizing I chose the COBYLA (Constrained Optimization BY Linear Approximations) algo, which allows inequality constraints without the need to specify gradients and jacobian matrices.



The video of me doing this live is here: http://vimeo.com/267043908

I managed to beat the optimizer.

# Initial test without copy/paste: 33333.83 seconds
```{r one}
# Set up objective function to mininimze
eval_f <- function(x) ( x[1] + x[2] +0.5*sign(x[2]) )
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] - 1e6))
# Initial guesses
x0 <- c(100,10000)
# Algo options (constrained optimization and tolerance)
opts <- list("algorithm"="NLOPT_LN_COBYLA",   "xtol_rel"=0.01)
# Run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0))
print(res)
```

# Copy/paste once: about 367 seconds
```{r two}
# Set up objective function to mininimze
eval_f <- function(x) ( x[1] + x[2]+0.5*sign(x[2]) + x[3]+1.5*sign(x[3]) )
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] + (5*x[1] + 30*x[2])*x[3] - 1e6))
# Initial guesses
x0 <- c(100,200,200)
# run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0,0))
print(res)
```

# Copy/paste twice: 99.5 seconds
```{r three}
# Set up 4-part objective function
eval_f <- function(x) ( x[1] + x[2]+0.5*sign(x[2]) + x[3]+1.5*sign(x[3]) + x[4]+1.5*sign(x[4]) )
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] + (5*x[1] + 30*x[2])*x[3] + 
                                    ((5*x[1] + 30*x[2])*x[3])*x[4] - 1e6))
# Initial guesses
x0 <- c(100,100,100,100)
# Run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0,0,0))
print(res)
# Round the results and check there are over 1e6 characters and total time
(32*30) + (32*30)*32 + (32*30)*32*32 + 0.5 +32 +1.5 +32 +1.5 +32
(0.5 +32) +(1.5 +32) +(1.5 +32)
```

# Copy/paste four times: 59 seconds
```{r four}
# Set up 5-part objective function
eval_f <- function(x) ( x[1] + x[2]+0.5*sign(x[2]) + x[3]+1.5*sign(x[3]) + x[4]+1.5*sign(x[4]) + x[5]+1.5*sign(x[5]) )
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] + (5*x[1] + 30*x[2])*x[3] + 
                                    ((5*x[1] + 30*x[2])*x[3])*x[4] +
                                  (((5*x[1] + 30*x[2])*x[3])*x[4])*x[5] - 1e6))
# Initial guesses
x0 <- c(0,32,32,32,32)
# Run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0,0,0,0))
print(res)
# Round the results and check there are over 1e6 characters and total time
(15*30) + (15*30)*13 + (15*30)*13*14 + (15*30)*13*14*12
(0.5 +15) +(1.5 +13) +(1.5 +14) +(1.5 +12)
```


# Copy/paste five times: 47.5 seconds
```{r five}
# Set up 6-part objective function
eval_f <- function(x) ( x[1] + x[2]+0.5*sign(x[2]) + x[3]+1.5*sign(x[3]) + x[4]+1.5*sign(x[4]) +  x[5]+1.5*sign(x[5]) +x[6]+1.5*sign(x[6]))
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] + (5*x[1] + 30*x[2])*x[3] + 
                                    ((5*x[1] + 30*x[2])*x[3])*x[4] +
                                    (((5*x[1] + 30*x[2])*x[3])*x[4])*x[5] +
                                  ((((5*x[1] + 30*x[2])*x[3])*x[4])*x[5])*x[6]- 1e6))
# Initial guesses
x0 <- c(0,15,15,15,15,15)
# Run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0,0,0,0,0))
print(res)
# Round the results and check there are over 1e6 characters and total time
(14*30) + (14*30)*7 + (14*30)*7*7 + (14*30)*7*7*7 + (14*30)*7*7*7*6
(0.5 +14) +(1.5 +7) +(1.5 +7) +(1.5 +7) +(1.5+6)
```



# Copy/paste six times: 41 seconds
```{r six}
# Set up 7-part objective function
eval_f <- function(x) ( x[1] + x[2]+0.5*sign(x[2]) + x[3]+1.5*sign(x[3]) + x[4]+1.5*sign(x[4]) + x[5]+1.5*sign(x[5]) +x[6]+1.5*sign(x[6]) +x[7]+1.5*sign(x[7]))
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] + (5*x[1] + 30*x[2])*x[3] + 
                                    ((5*x[1] + 30*x[2])*x[3])*x[4] +
                                    (((5*x[1] + 30*x[2])*x[3])*x[4])*x[5] +
                                    ((((5*x[1] + 30*x[2])*x[3])*x[4])*x[5])*x[6] +
                                    (((((5*x[1] + 30*x[2])*x[3])*x[4])*x[5])*x[6])*x[7] - 1e6))
# Initial guesses
x0 <- c(0,14,7,7,7,6,6)
# Run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0,0,0,0,0,0))
print(res)
# Round the results and check there are over 1e6 characters and total time
(6*30) + (6*30)*6 + (6*30)*6*6 + (6*30)*6*6*5 + (6*30)*6*6*5*5 + (6*30)*6*6*5*5*5
(0.5 +6) +(1.5 +6) +(1.5 +6) +(1.5 +5) +(1.5+5) +(1.5+5)
```


# Copy/paste seven times: 41.5 seconds > 41 seconds with six c/p cycles. Choose six cycles.
```{r seven}
# Set up 8-part objective functionv
eval_f <- function(x) ( x[1] + x[2]+0.5*sign(x[2]) + x[3]+1.5*sign(x[3]) + x[4]+1.5*sign(x[4]) + x[5]+1.5*sign(x[5]) +x[6]+1.5*sign(x[6]) +x[7]+1.5*sign(x[7]) +
                          +x[8]+1.5*sign(x[8]))
# Set up constraint function to 1e6 values
eval_g0 <- function(x)  return( -(5*x[1] + 30*x[2] + (5*x[1] + 30*x[2])*x[3] + 
                                    ((5*x[1] + 30*x[2])*x[3])*x[4] +
                                    (((5*x[1] + 30*x[2])*x[3])*x[4])*x[5] +
                                    ((((5*x[1] + 30*x[2])*x[3])*x[4])*x[5])*x[6] +
                                    (((((5*x[1] + 30*x[2])*x[3])*x[4])*x[5])*x[6])*x[7] +
                                    ((((((5*x[1] + 30*x[2])*x[3])*x[4])*x[5])*x[6])*x[7])*x[8]
                                    - 1e6))
# Initial guesses
x0 <- c(0,6,7,7,7,6,6,6)
# Run the minimization algo
res <- nloptr( x0=x0, eval_f=eval_f,eval_g_ineq = eval_g0, opts=opts,lb=c(0,0,0,0,0,0,0,0))
print(res)
# Round the results and check there are over 1e6 characters and total time
(5*30) + (5*30)*4 + (5*30)*4*5 + (5*30)*4*5*5 + (5*30)*4*5*5*4 + (5*30)*4*5*5*4*4 +(5*30)*4*5*5*4*4*3
(0.5 +6) +(1.5 +4) +(1.5 +5) +(1.5 +5) +(1.5+4) +(1.5+4) +(1.5+4)
```